Comparative Analysis of the wav2vec 2.0 Feature Extractor

### 2023-03-09 main.tex
add rasr/returnn, squeeze

\subsection{Training Details}
We generally follow the setup in \cite{zhou2022efficient}.
The batch size is 640k samples, which corresponds to \SI{40}{\second} of speech waveform and the gradients are accumulated over 3 steps.
The NAdam optimizer uses a \gls{OCLR} schedule with a peak learning rate of $3\cdot 10^{-4}$ and we train the model for 20 epochs.
In contrast to \cite{zhou2022efficient}, no gradient noise is applied.
Each model is trained on a single \gls{GPU}.

@@

The number of parameters for Gammatone and log Mel features refers to the \gls{FIR} and Mel filterbanks, respectively, that we use in our implementation even if they are not trainable.
The differences in the total number of parameters are influenced by the size of the \fe, but also of the linear layer before the \conformer which grows large as the feature dimension increases.
\input{tables/features_general}
\vspace{-0.3cm}

@@

\subsection{\wvtwo Pre-Training}
The main appeal of the \wvtwo framework is the ability to pre-train the model on unlabeled audio data.
\cite{facebook2020wav2vec2} even reports relative \gls{WER} improvements of about 10\% on the \textit{other} subsets for their experiments on LibriSpeech when pre-training on the same data as used in supervised training compared to pure supervised training from scratch.
Here, we study this effect only for the \fe by using the parameters of an \fe pre-trained on LibriSpeech\footnote{\raggedright\url{https://github.com/facebookresearch/fairseq/blob/main/examples/wav2vec/README.md}}, so no external data is added.
During the supervised training, the \fe weights can be further trained or kept frozen.
As shown in \refTab{table:features_pretraining}, no positive effect can be observed here.
In fact, using a frozen pre-trained \fe shows clearly worse performance than training the \fe purely supervised from scratch.
We hypothesize that the pre-training mainly benefits the \transformer encoder and not the \fe while the mismatch between the \transformer encoder in pre-training and \conformer encoder in our setup could also pose a challenge for the case that the \fe is frozen.
\input{tables/features_pretraining}

\subsection{Frequency Response of Learned Filters}
\label{sec:freq_resp}
To analyze the learned filters of the neural \fes, we first plot the frequency response of the filters in the first layer which operates on the raw waveform.
As the order of learned filters in a convolutional layer is arbitrary, we sort the filters by the peak value of the frequency response and by the upper and lower \SI{3}{\decibel} cutoff frequencies as second and third criteria.
The plots are depicted in \refFig{fig:first_layer}\,(a)-(d), which share the same color range.

@@

The distribution for the \wvtwo \fe's first layer in \refFig{fig:first_layer}\,(c) does not entail this degree of resemblance.
Again, bandpass filters are learned, however, they have a wider passband, their center frequencies are distributed rather linearly and the stopband attenuation is weak.
In addition, about 14\% of all filters have their peak at \SI{0}{\kilo\hertz} and another 10\% at \SI{8}{\kilo\hertz} with varying bandwidths.
Moreover, a number of filters exists with multiple passbands.
Interestingly, the frequency response of the pre-trained \wvtwo \fe in \refFig{fig:first_layer}\,(d) is very similar to \refFig{fig:first_layer}\,(c).
This indicates that the properties learned in unsupervised pre-training and supervised \gls{CTC} training are similar.

@@

\section{Acknowledgements}

\ifinterspeechfinal
     The authors would like to thank Wei Zhou for providing the baseline \gls{CTC} model.
\else
     The acknowledgements would likely reveal the author identity and are therefore hidden for the double-blind review process.
\fi

>>>>

\subsection{Training Details}
We generally follow the setup in \cite{zhou2022efficient}.
The batch size is 640k samples, which corresponds to \SI{40}{\second} of speech waveform and the gradients are accumulated over 3 steps.
The NAdam optimizer uses a \gls{OCLR} schedule with a peak learning rate of $3\cdot 10^{-4}$ and we train the model for 20 epochs.
In contrast to \cite{zhou2022efficient}, no gradient noise is applied.
Each model is trained on a single \gls{GPU}.
The experiments are carried out using 
\ifinterspeechfinal
     RETURNN and RASR
\else
     [\textit{toolkits blind for review}]
\fi
and configs will be made available at
\ifinterspeechfinal
     https://github.com/rwth-i6/.
\else
     [\textit{blind for review}].
\fi

@@

The number of parameters for Gammatone and log Mel features refers to the \gls{FIR} and Mel filterbanks, respectively, that we use in our implementation even if they are not trainable.
The differences in the total number of parameters are influenced by the size of the \fe, but also of the linear layer before the \conformer which grows large as the feature dimension increases.
\input{tables/features_general}

@@

\subsection{\wvtwo Pre-Training}
The main appeal of the \wvtwo framework is the ability to pre-train the model on unlabeled audio data.
\cite{facebook2020wav2vec2} even reports relative \gls{WER} improvements of about 10\% on the \textit{other} subsets for their experiments on LibriSpeech when pre-training on the same data as used in supervised training compared to pure supervised training from scratch.
Here, we study this effect only for the \fe by using the parameters of an \fe pre-trained on LibriSpeech\footnote{\raggedright\url{https://github.com/facebookresearch/fairseq/blob/main/examples/wav2vec/README.md}}, so no external data is added.
During the supervised training, the \fe weights can be further trained or kept frozen.
As shown in \refTab{table:features_pretraining}, no positive effect can be observed here.
In fact, using a frozen pre-trained \fe shows clearly worse performance than training the \fe purely supervised from scratch.
We hypothesize that the pre-training mainly benefits the \transformer encoder and not the \fe while the mismatch between the \transformer encoder in pre-training and \conformer encoder in our setup could also pose a challenge for the frozen \fe.
\input{tables/features_pretraining}

\vspace{-0.5cm}
\subsection{Frequency Response of Learned Filters}
\label{sec:freq_resp}
To analyze the learned filters of the neural \fes, we first plot the frequency response of the filters in the first layer which operates on the raw waveform.
As the order of learned filters in a convolutional layer is arbitrary, we sort the filters by the peak value of the frequency response and by the upper and lower \SI{3}{\decibel} cutoff frequencies as second and third criteria.
The plots are depicted in \refFig{fig:first_layer}\,(a)-(d), which share the same color range.

@@

The distribution for the \wvtwo \fe's first layer in \refFig{fig:first_layer}\,(c) does not entail this degree of resemblance.
Again, bandpass filters are learned, however, they have a wider passband, their center frequencies are distributed rather linearly and the stopband attenuation is weak.
In addition, about 14\% of all filters have their peak at \SI{0}{\kilo\hertz} and another 10\% at \SI{8}{\kilo\hertz} with varying bandwidths.
A number of filters exists with multiple passbands.
Interestingly, the frequency response of the pre-trained \wvtwo \fe in \refFig{fig:first_layer}\,(d) is very similar to \refFig{fig:first_layer}\,(c).
This indicates that the properties learned in unsupervised pre-training and supervised \gls{CTC} training are similar.

@@

\section{Acknowledgements}
\ifinterspeechfinal
     The authors would like to thank Wei Zhou for providing the baseline \gls{CTC} model.
\else
     The acknowledgements would likely reveal the author identity and are therefore hidden for the double-blind review process.
\fi
